Impact of Federal Party Leadership Changes on Donation Patterns

### 2024-12-03 paper/paper.qmd
update introduction

---
title: "Impact of Federal Party Leadership Changes on Donation Patterns"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

Overview paragraph

Estimand paragraph

Results paragraph

Why it matters paragraph

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....

>>>>

---
title: "Impact of Federal Party Leadership Changes on Donation Patterns"
subtitle: "An Analysis of Political Donations in Canada 2013 - 2024"
author: 
  - Maria Mangru
thanks: "Code and data are available at: [https://github.com/MariaMangru/Ontario_Political_Donors](https://github.com/MariaMangru/Ontario_Political_Donors)."
date: today
date-format: long
abstract: "This study investigates how the ruling political parties in Canada's federal and Ontario provincial governments affect the donation behaviors of individual contributors. By analyzing donation records from 2013 to 2024, the research explores whether a party being in power influences the total amount of financial support it receives. The findings reveal that opposition parties often receive more donations than those in power, highlighting donor preferences to support challengers. This insight enhances our understanding of political financing and can help parties and policymakers develop more effective fundraising strategies."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

Political donations play a critical role in supporting political campaigns and shaping electoral competition. In Canada, both federal and provincial elections rely heavily on contributions from individuals and organizations, which influence the political landscape and often reflect public support for different parties. Understanding what drives donation patterns is essential for ensuring transparency and fairness in the democratic process and for minimizing the risk of undue influence on policy-making.

One key question is whether a party’s position in power affects the amount of financial support it receives. While studies in other countries, such as the U.S.A, have shown that incumbents often benefit from more donations, particularly from interest groups looking to maintain influence, there has been less focus on whether this holds true in Canada. Research like Fouirnaies and Hall’s work (REFERENCE HERE) in the U.S. shows that donors strategically target those in power, but how these dynamics play out in Canadian elections remains under explored. This paper aims to fill that gap by analyzing how a party's power status influences donor behaviour in federal and Ontario provincial elections.

Using donation data from 2013 to 2024, this study compares the financial support received by parties in power versus those in opposition. The analysis examines total donations by year and party while accounting for factors like election timing and party size (major party or not). By identifying trends in how donors respond to shifts in political power, this research sheds light on the financial strategies behind Canada’s democratic processes.

The results show that opposition parties tend to receive higher donations than those in power, suggesting that donors often prioritize supporting challengers, possibly to influence political change. These findings align with trends observed in other political systems but highlight unique patterns in Canada. This information is valuable not only for political parties developing fundraising strategies but also for policymakers and researchers seeking to ensure fairness and accountability in political financing.

The paper is organized as follows: Section \ref{sec-data} discusses the data sources and methodology, followed by an explanation of the variables used in the analysis. The regression models and results are then presented, along with interpretations. The paper concludes by discussing the broader implications of the findings and suggesting directions for future research.[NEED TO UPDATE THE REFERENCES]
